Integral Transport Correction to Diffusion Calculations \\ Using the pCMFD Scheme

### 2020-01-05 docs/PHYSOR2020/ANE-RM-2020.tex
Update ANE-RM-2020.tex
Adding pCMFD equations.

\documentclass[review,3p,onecolumn,sort&compress]{elsarticle}
%
%  various packages that you may wish to activate for usage 
\usepackage{booktabs}
\usepackage{natbib}
%\usepackage{cite}
%\usepackage{epsf}
\usepackage{bm}
\usepackage[english]{babel}
\usepackage[autostyle, english = american]{csquotes}
\MakeOuterQuote{"}
\usepackage{color}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{calrsfs}
\usepackage{lineno}
\usepackage[colorlinks=true, a4paper=true, pdfstartview=FitV,
linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
\usepackage{xfrac}
\usepackage{units}
\usepackage{tikz}

@@

\newcommand{\eq}[1]{Eq.~(\ref{#1})}
\newcommand{\eqs}[1]{Eqs.~(\ref{#1})}
\newcommand{\tsup}[1]{\textsuperscript{#1}}
\newcommand{\tsub}[1]{\textsuberscript{#1}}
\newcommand{\keff}{\ensuremath{k_{\textrm{\scriptsize{eff}}}}}
\newcommand{\kinf}{\ensuremath{k_{\textrm{\scriptsize{$\infty$}}}}}
\newcommand{\pf}{\textrm{PPF}}
\newcommand{\eg}{e.g.,}
\newcommand{\ie}{i.e.,}
\newcommand{\OO}{\hat{\bm{\Omega}}}
\newcommand{\rr}{\bm{r}}
\newcommand{\jp}{\ensuremath{J^+}}
\newcommand{\jm}{\ensuremath{J^-}}
\newcommand{\jpm}{\ensuremath{J^\pm}}
\newcommand{\jD}{\ensuremath{J^{\textrm{\scriptsize{D}}}}}
\newcommand{\bmj}{\ensuremath{\bm{J}}}
\newcommand{\rE}{\ensuremath{(\rr,E)}}
\newcommand{\bnabla}{\ensuremath{\bm{\nabla}}}
\newcommand{\hzi}{\ensuremath{\sfrac{1}{2}}}
\newcommand{\mcL}{\tau}
% about notes and comments from each author
\newcommand{\erez}[1]{{\color{blue}{\textsuperscript{EREZ:}#1}}}
\newcommand{\roy}[1]{{\color{blue}{\textsuperscript{ROY:}#1}}}
\newcommand{\daniele}[1]{{\color{blue}{\textsuperscript{DANIELE:}#1}}}
\newcommand{\revised}[1]{{\color{red}{#1}}}

@@

		\title{INTEGRAL TRANSPORT CORRECTION TO DIFFUSION CALCULATIONS USING THE pCMFD SCHEME}

@@

\section{INTRODUCTION}
\label{sec:intro}

@@

% ------------------------------------------
\section{COARSE MESH FINITE DIFFERENCES}
\label{sec:CMFD-intro}

@@

% ------------------------------------------
\section{THE RONEN METHOD}
\label{sec:RM}

@@

% ------------------------------------------
\section{NUMERICAL IMPLEMENTATION}
\label{sec:RM-num}

@@

% ------------------------------------------
\subsection{CMFD IMPLEMENTATION}
\label{sec:RM-CMFD}

In classic CMFD implementation, the term $\delta J_{i+\hzi,g}$ is added to the diffusion current $\jD_{i+\hzi,g}$ in order to reproduce the net current $J_{i+\hzi,g}$. Currents are only computed on the surface elements of a coarse mesh in space to save memory and runtime \cite{Smith-1983}. Besides, the net current can be determined from some other expression which is physically or numerically more accurate. Therefore, this new term acts as a correction for the diffusion current. The diffusion current is generally approximated by finite differences according to
\begin{equation}
\label{eq:JD}
\jD_{i+\hzi,g} \cong - D_{i+\hzi,g}
\frac{\phi_{i+1,g} - \phi_{i,g}}{(\Delta_{i+1} + \Delta_i)/2},
\end{equation}
where the diffusion coefficient at the interface is volume-averaged as
\begin{equation}
\label{eq:Ds}
D_{i+\hzi,g} = \frac{\Delta_i D_{i,g} + \Delta_{i+1} D_{i+1,g}}
{\Delta_i + \Delta_{i+1}}.
\end{equation}
%
About the RM, more accurate estimates of the net current come from the integral expressions of \eqs{eq:par-curr-calc}, still using the scalar flux in the emission source $q_{i,g}$. In linearly anisotropic problems, the net current can also be considered to compute the source. Once the diffusion currents and the net currents by integral transport are calculated using the most recent flux values, the correction terms can be obtained on the cell interfaces. The discretized form of the terms $\delta J_{i+\hzi,g}$ reproduce a centered drift-advection term to avoid possible indeterminate division by zeros in case of flat flux~\cite{Smith-1983,Tomatis-2011}:
\begin{equation}
\label{eq:dJ}
\delta J_{i+\hzi,g} = -\delta D_{i+\hzi,g}
\frac{\phi_{i+1,g} + \phi_{i,g}}{(\Delta_{i+1} + \Delta_i)/2} =
J_{i+\hzi,g} - \jD_{i+\hzi,g},
\end{equation}
for $i = 0, \ldots, I-2$. The division by the spatial differences can be removed because the definition of $\delta D$ is arbitrary.

@@

% ------------------------------------------
\subsection{pCMFD IMPLEMENTATION}
\label{sec:RM-pCMFD}

@@

Two corrective terms are then introduced: $J = \jp - \jm = \jD + \delta J^+ + \delta J^-$. An upwind definition with respect to the direction of flight of neutrons is used for the corrections~\cite{Jarrett-2016,Zhu-2016}:
\begin{equation}
\label{eq:CCF-sub}
\delta \jp _{i+\hzi} = -\delta D^+ _{i+\hzi} \phi_i
\quad \text{and} \quad
\delta \jm _{i+\hzi} = -\delta D^- _{i+\hzi} \phi_{i+1}.
\end{equation}
The current by diffusion is distributed on the partial currents in equal parts in order to compute the new corrective coefficients $\delta D^{\pm}$,
\begin{subequations}
	\label{eq:delta-D-pCMFD}
	\begin{align}
	J^+_{i+\hzi} - \frac{1}{2}\jD_{i+\hzi} &= -\delta D^+_{i+\hzi} \phi_i, \label{eq:delta-D-pCMFDp}\\
	J^-_{i+\hzi} + \frac{1}{2}\jD_{i+\hzi} &= +\delta D^-_{i+\hzi} \phi_{i+1}, \label{eq:delta-D-pCMFDm}
	\end{align}
\end{subequations}
still for $i=0, \ldots, I-2$. The diffusion current is not halved at the boundary, using only \eq{eq:delta-D-pCMFDm} or \eq{eq:delta-D-pCMFDp} at the left and at the right boundaries, respectively. Finally, the implementation of CMFD can be upgraded to support the pCMFD scheme with minimal changes.
% -----------------------------------------------------------------------------
%
% Daniele: the following proposition has shown to be inconsistent.
%
% Thanks to P$_1$ diffusion theory, the partial currents at any point $x$ of the slab are approximated as
% \begin{equation}
% \label{eq:partial-current}
% \jpm(x) \cong \frac{\phi(x)}{4} \pm \frac{J (x)}{2},
% \end{equation}
% still verifying the expression for the net current $J = J^+ - J^-$. \eq{eq:partial-current} suggests introducing two different correction terms at each cell interface depending on the sign of the partial currents
% \begin{equation}
%   \label{eq:PCCF}
%   \jpm _{i+\hzi} = \frac{1}{4} \phi_{i+\hzi} \pm \frac{1}{2}
%     \left(\jD_{i+\hzi} + \delta \jpm_{i+\hzi}\right),
% \end{equation}
% and they must fulfill again the net current $J = \jD + (\delta \jp + \delta \jm)/2$. An upwind definition with the respect to the direction of flight of neutrons is~\cite{Jarrett-2016,Zhu-2016}
% \begin{equation}
% \label{eq:CCF-sub}
% \delta \jp _{i+\hzi} = -2\delta D^+ _{i+\hzi} \phi_i
% \quad \text{and} \quad
% \delta \jm _{i+\hzi} = -2\delta D^- _{i+\hzi} \phi_{i+1}.
% \end{equation}
% %
% Solving for the partial current correction factors gives
% \begin{equation}
%   \label{eq:ccfsss1}
%   \delta D^+ _{i+\hzi} = \frac{\frac{1}{4}\phi_{i+\hzi}
%   	+ \frac{1}{2}\jD _{i+\hzi} - \jp _{i+\hzi}}{\phi_i}
%   \quad \text{and} \quad
%   \delta D^- _{i+\hzi} = \frac{-\frac{1}{4}\phi_{i+\hzi}
%   	+ \frac{1}{2}\jD _{i+\hzi} + \jm _{i+\hzi}}{\phi_{i+1}},
%   \end{equation}
% where the flux at the interface can be simply volume-averaged too, see~\eq{eq:Ds} for instance.
% -----------------------------------------------------------------------------

@@

% ------------------------------------------
\section{RESULTS}
\label{sec:res}

% ------------------------------------------
\section{CONCLUSIONS}
\label{sec:conc}

>>>>

\documentclass[review,3p,onecolumn,sort&compress]{elsarticle}
%
%  various packages that you may wish to activate for usage 
\usepackage{booktabs}
\usepackage{natbib}
%\usepackage{cite}
%\usepackage{epsf}
\usepackage{bm}
\usepackage[english]{babel}
\usepackage[autostyle, english = american]{csquotes}
\MakeOuterQuote{"}
\usepackage{color}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{calrsfs}
\usepackage{lineno}
\usepackage[colorlinks=true, a4paper=true, pdfstartview=FitV,
linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}

% for algorithms
\usepackage{algorithm}
\usepackage{algpseudocode}

\usepackage{xfrac}
\usepackage{units}
\usepackage{tikz}

@@

\newcommand{\eq}[1]{Eq.~(\ref{#1})}
\newcommand{\eqs}[1]{Eqs.~(\ref{#1})}
\newcommand{\tsup}[1]{\textsuperscript{#1}}
\newcommand{\tsub}[1]{\textsuberscript{#1}}
\newcommand{\keff}{\ensuremath{k_{\textrm{\scriptsize{eff}}}}}
\newcommand{\kinf}{\ensuremath{k_{\textrm{\scriptsize{$\infty$}}}}}
\newcommand{\pf}{\textrm{PPF}}
\newcommand{\eg}{e.g.,}
\newcommand{\ie}{i.e.,}
\newcommand{\OO}{\hat{\bm{\Omega}}}
\newcommand{\rr}{\bm{r}}
\newcommand{\jp}{\ensuremath{J^+}}
\newcommand{\jm}{\ensuremath{J^-}}
\newcommand{\jpm}{\ensuremath{J^\pm}}
\newcommand{\delj}{\ensuremath{\delta J}}
\newcommand{\jD}{\ensuremath{J^{\textrm{\scriptsize{D}}}}}
\newcommand{\bmj}{\ensuremath{\bm{J}}}
\newcommand{\rE}{\ensuremath{(\rr,E)}}
\newcommand{\bnabla}{\ensuremath{\bm{\nabla}}}
\newcommand{\hzi}{\ensuremath{\sfrac{1}{2}}}
\newcommand{\mcL}{\tau}
\newcommand{\dx}{\Delta}
% about notes and comments from each author
\newcommand{\erez}[1]{{\color{blue}{\textsuperscript{EREZ:}#1}}}
\newcommand{\roy}[1]{{\color{blue}{\textsuperscript{ROY:}#1}}}
\newcommand{\daniele}[1]{{\color{blue}{\textsuperscript{DANIELE:}#1}}}
\newcommand{\revised}[1]{{\color{red}{#1}}}

@@

		\title{Integral Transport Correction to Diffusion Calculations \\ Using the pCMFD Scheme}

@@

\section{Intoduction}
\label{sec:intro}

@@

% ------------------------------------------
\section{Coarse Mesh Finite Differences}
\label{sec:CMFD-intro}

@@

% ------------------------------------------
\section{The Ronen Method}
\label{sec:RM}

@@

% ------------------------------------------
\section{Numerical Implementation}
\label{sec:RM-num}

@@

% ------------------------------------------
\subsection{CMFD Implementation}
\label{sec:RM-CMFD}

In classic CMFD implementation, the term $\delta J_{i+\hzi,g}$ is added to the diffusion current $\jD_{i+\hzi,g}$ in order to reproduce the net current $J_{i+\hzi,g}$. Currents are only computed on the surface elements of a coarse mesh in space to save memory and runtime \cite{Smith-1983}. Besides, the net current can be determined from some other expression which is physically or numerically more accurate. Therefore, this new term acts as a correction for the diffusion current. The diffusion current is generally approximated by discretizing Fick's law, and by finite differences, according to 
\begin{equation}
\label{eq:JD}
\jD_{i+\hzi,g} \cong - D_{i+\hzi,g}
\frac{\phi_{i+1,g} - \phi_{i,g}}{(\Delta_{i+1} + \Delta_i)/2},
\end{equation}
where the diffusion coefficient at the interface is volume-averaged as
\begin{equation}
\label{eq:Ds}
D_{i+\hzi,g} = \frac{\Delta_i D_{i,g} + \Delta_{i+1} D_{i+1,g}}
{\Delta_i + \Delta_{i+1}}.
\end{equation}
%
About the RM, more accurate estimates of the net current come from the integral expressions of \eq{eq:par-curr-calc}, still using the scalar flux in the emission source $q_{i,g}$. In linearly anisotropic problems, the net current can also be considered to compute the source. Once the diffusion currents and the net currents by integral transport are calculated using the most recent flux values, the correction terms can be obtained on the cell interfaces. The discretized form of the terms $\delta J_{i+\hzi,g}$ reproduce a centered drift-advection term to avoid possible indeterminate division by zeros in case of flat flux~\cite{Smith-1983,Tomatis-2011}:
\begin{equation}
\label{eq:dJ}
\delta J_{i+\hzi,g} = -\delta D_{i+\hzi,g}
\frac{\phi_{i+1,g} + \phi_{i,g}}{(\Delta_{i+1} + \Delta_i)/2} =
J_{i+\hzi,g} - \jD_{i+\hzi,g},
\end{equation}
for $i = 0, \ldots, I-2$. The division by the spatial differences can be removed because the definition of $\delta D$ is arbitrary.

@@

% ------------------------------------------
\subsection{pCMFD Implementation}
\label{sec:RM-pCMFD}

@@

Two corrective terms are then introduced: $J = \jp - \jm = \jD + \delta J^+ + \delta J^-$. An upwind definition with respect to the direction of flight of neutrons is used for the corrections~\cite{Jarrett-2016,Zhu-2016}:
\begin{equation}
\label{eq:CCF-sub}
\delta \jp _{i+\hzi} = -\delta D^+ _{i+\hzi} \phi_i
\quad \text{and} \quad
\delta \jm _{i+\hzi} = -\delta D^- _{i+\hzi} \phi_{i+1}.
\end{equation}
The current by diffusion is distributed on the partial currents in equal parts in order to compute the new corrective coefficients $\delta D^{\pm}$,
\begin{subequations}
	\label{eq:delta-D-pCMFD}
	\begin{align}
	J^+_{i+\hzi} - \frac{1}{2}\jD_{i+\hzi} &= -\delta D^+_{i+\hzi} \phi_i, \label{eq:delta-D-pCMFDp}\\
	J^-_{i+\hzi} + \frac{1}{2}\jD_{i+\hzi} &= +\delta D^-_{i+\hzi} \phi_{i+1}, \label{eq:delta-D-pCMFDm}
	\end{align}
\end{subequations}
still for $i=0, \ldots, I-2$. The diffusion current is not halved at the boundary, using only \eq{eq:delta-D-pCMFDm} or \eq{eq:delta-D-pCMFDp} at the left and at the right boundaries, respectively. Finally, the implementation of CMFD can be upgraded to support the pCMFD scheme with minimal changes.
% -----------------------------------------------------------------------------
%
% Daniele: the following proposition has shown to be inconsistent.
%
% Thanks to P$_1$ diffusion theory, the partial currents at any point $x$ of the slab are approximated as
% \begin{equation}
% \label{eq:partial-current}
% \jpm(x) \cong \frac{\phi(x)}{4} \pm \frac{J (x)}{2},
% \end{equation}
% still verifying the expression for the net current $J = J^+ - J^-$. \eq{eq:partial-current} suggests introducing two different correction terms at each cell interface depending on the sign of the partial currents
% \begin{equation}
%   \label{eq:PCCF}
%   \jpm _{i+\hzi} = \frac{1}{4} \phi_{i+\hzi} \pm \frac{1}{2}
%     \left(\jD_{i+\hzi} + \delta \jpm_{i+\hzi}\right),
% \end{equation}
% and they must fulfill again the net current $J = \jD + (\delta \jp + \delta \jm)/2$. An upwind definition with the respect to the direction of flight of neutrons is~\cite{Jarrett-2016,Zhu-2016}
% \begin{equation}
% \label{eq:CCF-sub}
% \delta \jp _{i+\hzi} = -2\delta D^+ _{i+\hzi} \phi_i
% \quad \text{and} \quad
% \delta \jm _{i+\hzi} = -2\delta D^- _{i+\hzi} \phi_{i+1}.
% \end{equation}
% %
% Solving for the partial current correction factors gives
% \begin{equation}
%   \label{eq:ccfsss1}
%   \delta D^+ _{i+\hzi} = \frac{\frac{1}{4}\phi_{i+\hzi}
%   	+ \frac{1}{2}\jD _{i+\hzi} - \jp _{i+\hzi}}{\phi_i}
%   \quad \text{and} \quad
%   \delta D^- _{i+\hzi} = \frac{-\frac{1}{4}\phi_{i+\hzi}
%   	+ \frac{1}{2}\jD _{i+\hzi} + \jm _{i+\hzi}}{\phi_{i+1}},
%   \end{equation}
% where the flux at the interface can be simply volume-averaged too, see~\eq{eq:Ds} for instance.
% -----------------------------------------------------------------------------
The descretized form of the diffusion equations, with pCMFD implementations, is
\begin{equation}
	\label{eq:pCMFD-diff1}
	\begin{split}
	J_{D,i+\hzi,g}^+ +\left( \delj_{i+\hzi,g}^+ - \delj_{i+\hzi,g}^- \right) - J_{D,i-\hzi,g}^+ &+ \left( + \delj_{i-\hzi,g}^+ - 
	\delj_{i-\hzi,g}^- \right) + \sigma_{i,g}\phi_{i,g} = \\
	&\frac{\chi_g}{\keff}\sum\limits^G_{g'=1}\nu\sigma_{f,g'}\phi_{i,g'} + \sum\limits_{g'=1}^G \sigma_{s,i,0,g\leftarrow g'}\phi_{i,g'}\,\,.
	\end{split}
\end{equation}
By using \eqs{eq:CCF-sub}, one gets
\begin{equation}
	\label{eq:pCMFD-diff2}
	-2D_{i+\hzi}\frac{\phi_{i+1,g} - \phi{i,g}}{\dx_{i+1}+\dx_{i}} 
\end{equation}

@@

% ------------------------------------------
\section{Results}
\label{sec:res}

% ------------------------------------------
\section{Conclusions}
\label{sec:conc}
